---
title: "Covid_19_Project"
author: "Andrew Irani"
date: "9/22/2020"
output: html_document
---

A pneumonia of unknown cause detected in Wuhan, China was first internationally reported from China on 31 December 2019. Today we know this virus as Coronavirus. COVID-19 which stands for COronaVIrus Disease is the disease caused by this virus. Since then, the world has been engaged in the fight against this pandemic. Several measures have therefore been taken to "flatten the curve". We have consequently experienced social distancing and many people have passed away as well.

In the solidarity to face this unprecedented global crisis, several organizations did not hesitate to share several datasets allowing the conduction of several kinds of analysis in order to understand this pandemic.

It is natural for us to analyze these datasets by ourselves to answer questions since we cannot always rely on the news, and we are data scientists.

This analysis tries to provide an answer to this question: **Which countries have had the highest number of positive cases against the number of tests?**

# Step 1: Understanding the data
```{r}
#install.packages('tidyverse')
library(tidyverse)
```

```{r}
# Load the COVID dataset
covid_df <- read.csv("covid19.csv")
```

```{r}
# Data dimensions
library(tibble)
dim(covid_df)
glimpse(covid_df)
```

```{r}
# Display variables
vector_cols <- colnames(covid_df)
vector_cols
```

```{r}
# Quick overview of the data
head(covid_df)
glimpse(covid_df)
```

# Step 2: Cleaning up data
```{r}
# Extract only country level data
covid_df_all_states <- covid_df %>% 
  filter(Province_State == "All States") %>% 
         select(-Province_State)
# After filtering for "All states" the Province_State column is no longer needed because that's all it contains
```

```{r}
# Select the columns that are daily
covid_df_all_states_daily <- covid_df_all_states %>% 
  select(Date, Country_Region, active, hospitalizedCurr, daily_tested, daily_positive)

```

# Step 3: Summarize the data
```{r}
# For each country get the total number of tested, positive, active, and hosipitalized
covid_df_all_states_daily_sum <- covid_df_all_states_daily %>% 
  group_by(Country_Region) %>% 
  summarise(tested_sum = sum(daily_tested),
            positive_sum = sum(daily_positive),
            active_sum = sum(active),
            hospitalized_sum = sum(hospitalizedCurr)) %>% 
              arrange(-tested_sum)
covid_df_all_states_daily_sum
            
            
            
```

```{r}
# Store the top 10 countries
covid_top_10 <- head(covid_df_all_states_daily_sum, 10)

```

# Step 4: Question = Which countries have had the highest number of postive cases against the number of tests
```{r}
# Answer the question using vectors
countries <- covid_top_10$Country_Region
tested_cases <- covid_top_10$tested_sum
positive_cases <- covid_top_10$positive_sum
active_cases <- covid_top_10$active_sum
hospitalized_cases <- covid_top_10$hospitalized_sum

```

```{r}
# Naming vectors
names(positive_cases) <- countries
names(tested_cases) <- countries
names(active_cases) <- countries
names(hospitalized_cases) <- countries
```

```{r}
# Identify the top 3 countries
positive_tested_division <- positive_cases/tested_cases
positive_tested_division_sorted <- sort(positive_tested_division, decreasing = TRUE)
positive_tested_division_sorted

```

```{r}
# Display the top 3 countries
positive_tested_top_3 <- c("United Kingdom = 0.11", "United States = 0.10", "Turkey = 0.08")

```

# Step 5: Complete the previous step using Tidyverse Dplyr
```{r}
# Same answer can be created using more robust code
positive_tested_top_3_tidy <- covid_top_10 %>% 
  group_by(Country_Region) %>% 
  mutate(positive_vs_tested = positive_sum/tested_sum) %>% 
  arrange(-positive_vs_tested) %>% 
  select(positive_vs_tested, Country_Region) %>%
  head(3)
positive_tested_top_3_tidy


```

# Step 6: Create a matrix using the information from vectors
```{r}
# Create new vectors
united_kingdom <- c(0.11, 1473672, 166909, 0, 0)
united_states <- c(0.10, 17282363, 1877179, 0, 0)
turkey <- c(0.08, 2031192, 163941, 2980960, 0)

```

```{r}
# Create and display matrix using the vectors
covid_mat <- rbind(united_kingdom, united_states, turkey)
colnames(covid_mat) <- c("Ratio", "tested", "positive", "active", "hospitalized")
covid_mat

```


# Step 7: Putting everything together using lists
```{r}
# Create question variable
question <- "Which countries have had the highest number of positive cases against the number of tests?"

# Create answer variable
answer <- c("Positive tested cases" = positive_tested_top_3)

```

```{r}
# Create a list that contains previous data structures
dataframe <- list(covid_df, covid_df_all_states, covid_df_all_states_daily, covid_top_10)
matrices <- list(covid_mat)
vectors <- list(vector_cols, countries)
data_structure_list <- list("dataframe" = dataframe, "matrix" = matrices, "vector" = vectors)
covid_analysis_list <- list(question, answer, data_structure_list)

```

```{r}
# Display the entire list
covid_analysis_list[[2]]
```

# Step 8: Conclusion

We can see the top 3 countries with the highest ratio of positive tests to the number of tests are the United Kingdom, United States and Turkey.
<br>
The project mainly focused on using the techniques learned in the section including vectors, lists and matrices. However all of these steps could have been done more efficiently with Dplyr. So, going forward I'll focus more on learning the tidyverse because it seems so intuitive.




