---
title: "Covid_19_Project: Continued Analysis"
author: "Andrew Irani"
date: "9/22/2020"
output: html_document
---

A pneumonia of unknown cause detected in Wuhan, China was first internationally reported from China on 31 December 2019. Today we know this virus as Coronavirus. COVID-19 which stands for COronaVIrus Disease is the disease caused by this virus. Since then, the world has been engaged in the fight against this pandemic. Several measures have therefore been taken to "flatten the curve". We have consequently experienced social distancing and many people have passed away as well.

In the solidarity to face this unprecedented global crisis, several organizations did not hesitate to share several datasets allowing the conduction of several kinds of analysis in order to understand this pandemic.

It is natural for us to analyze these datasets by ourselves to answer questions since we cannot always rely on the news, and we are data scientists.

This analysis tries to provide an answer to this question: **Which countries have had the highest number of positive cases against the number of tests?**

# Step 1: Understanding the data
```{r}
#install.packages('tidyverse')
library(tidyverse)
```

```{r}
# Load the COVID dataset
covid_df <- read.csv("covid19.csv")
```

```{r}
# Data dimensions
library(tibble)
dim(covid_df)
glimpse(covid_df)
```

```{r}
# Display variables
vector_cols <- colnames(covid_df)
vector_cols
```

```{r}
# Quick overview of the data
head(covid_df)
glimpse(covid_df)
```

# Step 2: Cleaning up data
```{r}
# Extract only country level data
covid_df_all_states <- covid_df %>% 
  filter(Province_State == "All States") %>% 
         select(-Province_State)

# After filtering for "All states" the Province_State column is no longer needed because that's all it contains
```

```{r}
# Select the columns that are daily
covid_df_all_states_daily <- covid_df_all_states %>% 
  select(Date, Country_Region, active, hospitalizedCurr, daily_tested, daily_positive)

# Select the columns that are related to cumulative measures 
covid_df_all_states_cumulative <- covid_df_all_states %>% 
  select(Date, Continent_Name, Two_Letter_Country_Code, positive, hospitalized, recovered, death, total_tested)

```

# Step 3: Question - Which countries have had the highest number of deaths due to COVID-19
```{r}
# Summarize the dataset
covid_df_all_states_cumulative_max <- covid_df_all_states_cumulative %>% 
  group_by(Continent_Name, Two_Letter_Country_Code) %>% 
summarise(death_max = max(death)) %>% 
  filter(death_max > 0) %>% 
  arrange(-death_max)

# Display the dataset
covid_df_all_states_cumulative_max

```

```{r}
# Visualize the maximum number of deaths for each country
death_top_3 <-  qplot(x = Two_Letter_Country_Code,
                      y = death_max,
                      color = Continent_Name,
                      data = covid_df_all_states_cumulative_max)
death_top_3

# This graph works but is quite cluttered on the x axis
```

```{r}
# Create a graph that is easier to interpret
death_top_3 <-  qplot(x = death_max,
                      y = Two_Letter_Country_Code,
                      color = Continent_Name,
                      data = covid_df_all_states_cumulative_max)
death_top_3 

# Just buy switching the x and y it makes the entire graph more useful
```

```{r}
# Store the top 3 countris
death_top_3 <- c("US", "Italy", "GB")
```

# Step 4: Summarize the data
```{r}
# For each country get the total number of tested, positive, active, and hosipitalized
covid_df_all_states_daily_sum <- covid_df_all_states_daily %>% 
  group_by(Country_Region) %>% 
  summarise(tested_sum = sum(daily_tested),
            positive_sum = sum(daily_positive),
            active_sum = sum(active),
            hospitalized_sum = sum(hospitalizedCurr)) %>% 
              arrange(-tested_sum)
head(covid_df_all_states_daily_sum)
            
            
            
```

```{r}
# Store the top 10 countries
covid_top_10 <- head(covid_df_all_states_daily_sum, 10)

```

# Step 5: Question = Which countries have had the highest number of postive cases against the number of tests
```{r}
# Answer the question using vectors
countries <- covid_top_10$Country_Region
tested_cases <- covid_top_10$tested_sum
positive_cases <- covid_top_10$positive_sum
active_cases <- covid_top_10$active_sum
hospitalized_cases <- covid_top_10$hospitalized_sum

```

```{r}
# Naming vectors
names(positive_cases) <- countries
names(tested_cases) <- countries
names(active_cases) <- countries
names(hospitalized_cases) <- countries
```

```{r}
# Identify the top 3 countries
positive_tested_division <- positive_cases/tested_cases
positive_tested_division_sorted <- sort(positive_tested_division, decreasing = TRUE)
positive_tested_division_sorted

```

```{r}
# Display the top 3 countries
positive_tested_top_3 <- c("United Kingdom" = 0.11, "United States" = 0.10, "Turkey" = 0.08)

```

# Step 6: Complete the previous step using Tidyverse Dplyr
```{r}
# Same answer can be created using more robust code
positive_tested_top_3_tidy <- covid_top_10 %>% 
  group_by(Country_Region) %>% 
  mutate(positive_vs_tested = positive_sum/tested_sum) %>% 
  arrange(-positive_vs_tested) %>% 
  select(positive_vs_tested, Country_Region) %>%
  head(3)
positive_tested_top_3_tidy


```

# Step 7: Quantify the relationship between the number of cases and the population of each country
```{r}
# Create a matrix using the previous vectors
covid_mat <- cbind(tested_cases, positive_cases, active_cases, hospitalized_cases)

# Create a population vector
population <- c(331002651, 145934462, 60461826, 1380004385, 84339067, 37742154, 67886011, 25499884, 32971854, 37846611)

# complete the matrix
covid_mat <- covid_mat * 100/population

# Display the matrix
covid_mat

```

```{r}
covid_mat_tested_cases <- covid_mat[, "tested_cases"]
                      
covid_mat_tested_cases
```



# Step 8: Analyze how countries have been affected relative to each other using rankings
```{r}
# Create ranking vectors
tested_cases_rank <- rank(covid_mat[, "tested_cases"])
positive_cases_rank <- rank(covid_mat[, "positive_cases"])
active_cases_rank <- rank(covid_mat[, "active_cases"])
hospitalized_cases_rank <- rank(covid_mat[, "hospitalized_cases"])

# Create matrix combining the ranking vectors
covid_mat_rank <- rbind(tested_cases_rank, positive_cases_rank, active_cases_rank, hospitalized_cases_rank)

# Display the matrx
covid_mat_rank

# Display the first row of the matrix
covid_mat_rank[1,]

# Remove the first row of the matrix
covid_mat_rank[-1, ]

# Compute the sum of the rows
colSums(covid_mat_rank[-1, ])
                        
                                
```

# Step 9: Questions- 
*Which countries have made the best effort in terms of the number of tests conducted related to their population?

*Which countries were ultimately the most and least affected related to their population?
```{r}
# Identify the top 3 countries that made the best effort in terms of the number of tests conducted relative to their population
best_effort_tested_cased_top_3 <- c("India" = 1.0,
                                    "United Kingdom" = 2.0, 
                                    "Turkey" = 3.0)

# Most affected country 
most_affected_country <- ("Italy")

# Least affected country
least_affected_country <- ("India")

```

# Step 10: Combine all datasets and answers into one readable lists that summarizes the project
```{r}
# Create a list of all the previous questions
question_list <- list( "Which countries have had the highest fatality (mortality) rates?",
  "Which countries have had the highest number of positive cases against the number of tests?",
  "Which countries have made the best effort in terms of the number of tests conducted related to their population?",
  "Which countries were ultimately the most and least affected related to their population?"
)

# Create a list that contains all the answers
answer_list <- list("Death" = 
                      death_top_3,
                    "Positive tested cases" = 
                      positive_tested_top_3,
                    "The best effort in test related to the population" =
                      best_effort_tested_cased_top_3,
                    "The most affected country related to its population" =
                      most_affected_country,
                    "The least affected country related to its population" =
                      least_affected_country)

# Display answer list
answer_list

# Create a list containing all of the dataframes
dataframes <- list(covid_df, 
                   covid_df_all_states, 
                   covid_df_all_states_cumulative, 
                   covid_df_all_states_daily)


# Create a list containing all of the matrices
matrices <- list(covid_mat, 
                 covid_mat_rank)

# Create a list containing all of the vectors
vectors <- list(vector_cols,
                population,
                countries)

# Create a list that contains the 3 previous lists
data_structure_list <- list("data frame" = dataframes,
                            "matrix" = matrices,
                            "vector" = vectors)

# Create a list that contains all previously made lists
covid_analysis_list <- list(question_list, answer_list, data_structure_list)

```

# Step 11: Conclusion

The top 3 countries with the highest ratio of positive tests to the number of tests are the United Kingdom, United States and Turkey.
<br>
The project mainly focused on using the techniques learned in the section including vectors, lists and matrices. However all of these steps could have been done more efficiently with Dplyr. So, going forward I'll focus more on learning the tidyverse because it seems so intuitive.
<br>

# Step 12: Continued Conclusion

The countries that have given the best effort to fighting COVID-19 relative to population are India, United Kingdom and Turkey.
<br>
The most affected country relative to population is Italy
<br>
The least affected country relative to population is India
<br>
This project was a more in depth analysis of the previous COVID-19 project with more emphasis on using Dplyr and graphing techniques to answer deeper questions.
However the code could definitetly be refined more using the tidyverse packages better. Overall the code is becoming more understandable and I'm picking the logic much quicker. Also typing speed has greatly increased with fewer mistakes. Another interesting note is that my commenting, spacing and styling of the code has become much more clear. I found the provided solution code to be quite clunky, unclear and not commented well.













